### 2021-07-08 ch/1-introduction.tex
Start writing the new sections

Now, one possible objection here is:
if those datasets are of such poor quality,
then maybe we just should not train any \ac{ML} model on these and should not use them to make automated decisions?
While this question is mostly beyond the scope of this document, let me offer some thoughts on this:
It is true that even after the application of de-biasing techniques,
the resulting models still should not be fully trusted,
but they can still ease the burden of checking everything manually;
similar to an email spam filter which is not perfect, but still very useful.
Or, put another way, it is always important to check what the realistic alternative is;
we should not compare a model to a non-existing perfect ideal, but to the actual solution that would be used instead.
One could imagine a hybrid approach where an automated system makes preliminary decisions,
but random samples are reviewed by humans and decisions can always be challenged.
Ideally, the model itself would tell us about decisions it is uncertain about.

Moreover, two of the three methods presented in this thesis require access to (unlabelled) \emph{unbiased} data,
which is used during the training process.
The remaining method requires summary statistics of the unbiased target dataset.
So, the criticism that we are only learning from biased data does not apply here.
This should allow us to be more confident in the predictions produced by those methods,
because the methods learn from additional, unbiased information.

@@

% \section{Claims and contributions}%
\section{List of publications and author contributions}%
\label{sec:claims-contributions}
This thesis is based on 3 publications, corresponding to \rangechapref{ch:paper1}{ch:paper3}.
The first one is concerned with \emph{label bias} and the other two with \emph{sampling bias}.

\subsection{Publication 1}
\begin{refsection}[allreferences]
    \nocite{kehrenberg2020tuning}
    \printbibliography[heading=none]
\end{refsection}
\noindent A shorter version was published as a workshop paper:
\begin{refsection}[allreferences]
    \nocite{kehrenberg2018interpretable}
    \printbibliography[heading=none]
\end{refsection}
\noindent\textsc{Contributions:}
\begin{itemize}
  \item I conceived the idea of using Target Labels to target a balanced set. I developed the proof from a starting point that my supervisor pointed me to. I wrote all of the code dealing with the modified loss function and nearly all of the remaining code as well. I ran most of the experiments and wrote all of the methods section and most of the remaining text as well.
  \item Z. Chen was a discussion partner and helped run the experiments, and wrote some parts of the code.
  \item N. Quadrianto suggested the initial direction of the work, was a discussion partner, and helped write the introduction and related work.
\end{itemize}

\subsection{Publication 2}
\begin{refsection}[allreferences]
    \nocite{kehrenberg2020nullsampling}
    \printbibliography[heading=none]
\end{refsection}%
\noindent\textsc{Contributions:}
\begin{itemize}
  \item I conceived the idea of using an \acf{INN} and a representative set to learn an invariant representation. I wrote a large part of the code, and ran about half the experiments. I wrote a significant part of the text.
  \item M. Bartlett developed a large part of the tricks needed for training the \ac{INN} successfully. He wrote a significant part of the code, and of the text.
  \item O. Thomas helped with writing the code and with running the experiments.
  \item N. Quadrianto gave feedback on the progress and suggested directions to explore.
\end{itemize}

\subsection{Publication 3}
\begin{refsection}[allreferences]
    \nocite{kehrenberg2020zeroshot}
    \printbibliography[heading=none]
\end{refsection}%
\noindent\textsc{Contributions:}
\begin{itemize}
  \item I developed the idea of distribution matching as an extension of the previous paper. In addition, I tried to achieve similar goals by applying clustering to an unlabelled auxiliary set, with supervision from the labelled training set. I wrote all of the initial implementation of the method, and a large part of the later refined implementation. I ran the majority of the experiments.
  \item V. Sharmanska developed the original link to a fairness problem and wrote parts of the introduction and related work and contributed to other sections.
  \item M. Bartlett introduced the idea of batches-of-bags. He also improved the \ac{NN} architectures of the encoder and the discriminator, ran experiments, wrote the sections on architecture and contributed to other parts of the paper.
  \item N. Quadrianto suggested how to combine the two research directions that I had into a coherent whole. He also gave general feedback and suggested directions to explore.
\end{itemize}

\section{Structure of document}%
\label{sec:thesis-structure}
In \chapref{ch:related-work}, I present related work,
concentrating on the area of algorithmic fairness and dataset bias.
\Chapref{ch:content} summarises the main contributions of this thesis
and describes their relationship to one another.
\Rangechapref{ch:paper1}{ch:paper3} reproduce the publications listed in the previous section.
% constitute the main contribution of this thesis.
% Each of these chapters starts with a brief section which positions the chapter in the context of the this document.
% This is followed by the contents of publications.
In the final chapter, \chapref{ch:conclusion}, I present conclusions from the main work and directions for future work.

>>>>

% Now, one possible objection here is:
% if those datasets are of such poor quality,
% then maybe we just should not train any \ac{ML} model on these and should not use them to make automated decisions?
% While this question is mostly beyond the scope of this document, let me offer some thoughts on this:
% It is true that even after the application of de-biasing techniques,
% the resulting models still should not be fully trusted,
% but they can still ease the burden of checking everything manually;
% similar to an email spam filter which is not perfect, but still very useful.
% Or, put another way, it is always important to check what the realistic alternative is;
% we should not compare a model to a non-existing perfect ideal, but to the actual solution that would be used instead.
% One could imagine a hybrid approach where an automated system makes preliminary decisions,
% but random samples are reviewed by humans and decisions can always be challenged.
% Ideally, the model itself would tell us about decisions it is uncertain about.

% Moreover, two of the three methods presented in this thesis require access to (unlabelled) \emph{unbiased} data,
% which is used during the training process.
% The remaining method requires summary statistics of the unbiased target dataset.
% So, the criticism that we are only learning from biased data does not apply here.
% This should allow us to be more confident in the predictions produced by those methods,
% because the methods learn from additional, unbiased information.

@@

\section{Connection to other fields}
The thesis is mostly written from the perspective of algorithmic  fairness, but it touches on other fields as well, like \emph{domain shift} and \emph{causality}.
A domain shift is, in general, a change in the data distribution between the training set and the deployment setting;
one often talks of a ``source domain'' and a ``target domain''.
Typically, this leads to poor performance of the \ac{ML} system in the deployment setting,
which necessitates the development of \emph{domain adaptation} methods.
There are thus strong parallels to the problem of dataset bias and fairness,
but domain shift is more general and has a different emphasis.
In domain shift, it is the whole data distribution that changes, including the meaning of individual features,
whereas in the biased-data setting (as defined in this thesis), the distribution of training data broadly matches that of the deployment setting, except for incorrect labels or censored sampling.
Furthermore, the field of algorithmic fairness is characterised by a focus on special attributes (sensitive attributes) that should not be used to make predictions;
a focus that the problem of domain shift lacks.
Nevertheless, methods developed for domain adaptation can often be adapted to the problem of biased data:
A good amount of the prior work discussed in \chapref{ch:related-work} was motivated by domain adaptation instead of fairness.
Similarly, it should be possible to adapt fairness methods to tackle domain shift,
however, this has rarely happened.

The other closely related topic is \emph{causality}.
The argument that a solution to the dataset bias problem involves causality goes as follows:
As mentioned above, the goal is to ignore the bias in the data
and learn the true underlying structure that is hidden in the data.
Often, the most fundamental structure that we can learn is the \emph{causal} structure of the problem of interest.
For example, if an \acs{ML} model truly understood what makes someone a good employee,
it would not need to rely on surface characteristics like whether the applicants name sounds foreign.
% Learning such causal models is not yet completely in the realm of what today's \ac{ML} systems can achieve \citep{pearl2019seven},
% and will likely require a very large amount of data.
% (The models that perhaps come close to this are methods like unsupervised methods like GPT \citep{radford2018improving,radford2019language,brown2020language} that have been trained on enormous amounts of data.)
However, the work in this thesis is not presented under the banner of causality, for several reasons.
% First, a majority of the experiments are on images, such that the input features are pixel values,
% and thus do not lend themselves well to be analysed with methods like Bayesian
First, in as far as the presented methods solve a causal structure problem, they solve a very limited one;
inputs are mapped to outputs, and no detailed exploration of a potential implicit causal model is performed.
Second, the discovered robust structures are not necessarily \emph{causal} in nature, in a narrow sense of the word.
This is especially true in image data:
For example, while discovering that the presence of a smile and the gender of the person in a photograph are two distinct characteristics
requires a deep understanding of human faces,
it does not require knowing how smiles and gender are related via cause-and-effect in the real world.

In the chapter on related work (\chapref{ch:related-work}),
I discuss several methods that apply proper causal methods to fairness,
but those methods assume that the true causal structure of the problem is already known,
and are not helpful for discovering such structures.
To work properly, they need access to either Bayesian networks or structural equation models.
Neither of which the methods presented in this thesis can provide.

Finally, I will mention transfer learning and terms like interpretability and explainability.

\section{Structure of document}%
\label{sec:thesis-structure}
In \chapref{ch:related-work}, I present related work,
concentrating on the area of algorithmic fairness and dataset bias.
\Chapref{ch:content} summarises the main contributions of this thesis
and describes their relationship to one another.
The chapter concludes with a list of the publications that constitute the main contribution of the thesis,
accompanied by a description of the contributions to the publications, separated by author.
\Rangechapref{ch:paper1}{ch:paper3} reproduce these publications with minimal changes.
% constitute the main contribution of this thesis.
% Each of these chapters starts with a brief section which positions the chapter in the context of the this document.
% This is followed by the contents of publications.
In the final chapter, \chapref{ch:conclusion}, I present conclusions from the main work and directions for future work.

### 2021-07-08 ch/3-structure.tex
Start writing the new sections

>>>>

% \section{Claims and contributions}%
\section{List of publications and author contributions}%
\label{sec:claims-contributions}
This thesis is based on 3 publications, corresponding to \rangechapref{ch:paper1}{ch:paper3}.
The first one is concerned with \emph{label bias} and the other two with \emph{sampling bias}.

\subsection{Publication 1}
\begin{refsection}[allreferences]
    \nocite{kehrenberg2020tuning}
    \printbibliography[heading=none]
\end{refsection}
\noindent A shorter version was published as a workshop paper:
\begin{refsection}[allreferences]
    \nocite{kehrenberg2018interpretable}
    \printbibliography[heading=none]
\end{refsection}
\noindent\textsc{Contributions:}
\begin{itemize}
  \item I conceived the idea of using Target Labels to target a balanced set.
    I developed the proof from a starting point that my supervisor pointed me to.
    I wrote all of the code dealing with the modified loss function and nearly all of the remaining code as well.
    I ran most of the experiments and wrote all of the methods section and most of the remaining text as well.
  \item Z. Chen was a discussion partner and helped run the experiments, and wrote some parts of the code.
  \item N. Quadrianto suggested the initial direction of the work, was a discussion partner,
    and helped write the introduction and related work.
\end{itemize}

\subsection{Publication 2}
\begin{refsection}[allreferences]
    \nocite{kehrenberg2020nullsampling}
    \printbibliography[heading=none]
\end{refsection}%
\noindent\textsc{Contributions:}
\begin{itemize}
  \item I conceived the idea of using an \acf{INN} and a representative set to learn an invariant representation.
    I wrote a large part of the code, and ran about half the experiments. I wrote a significant part of the text.
  \item M. Bartlett developed a large part of the tricks needed for training the \ac{INN} successfully.
    He wrote a significant part of the code, and of the text.
  \item O. Thomas helped with writing the code and with running the experiments.
  \item N. Quadrianto gave feedback on the progress and suggested directions to explore.
\end{itemize}

\subsection{Publication 3}
\begin{refsection}[allreferences]
    \nocite{kehrenberg2020zeroshot}
    \printbibliography[heading=none]
\end{refsection}%
\noindent\textsc{Contributions:}
\begin{itemize}
  \item I developed the idea of distribution matching as an extension of the previous paper.
    In addition, I tried to achieve similar goals by applying clustering to an unlabelled auxiliary set,
    with supervision from the labelled training set.
    I wrote all of the initial implementation of the method, and a large part of the later refined implementation.
    I ran the majority of the experiments.
  \item V. Sharmanska developed the original link to a fairness problem
    and wrote parts of the introduction and related work and contributed to other sections.
  \item M. Bartlett introduced the idea of batches-of-bags.
    He also improved the \ac{NN} architectures of the encoder and the discriminator,
    ran experiments, wrote the sections on architecture and contributed to other parts of the paper.
  \item N. Quadrianto suggested how to combine the two research directions that I had into a coherent whole.
    He also gave general feedback and suggested directions to explore.
\end{itemize}

### 2021-07-08 ch/7-discussion.tex
Start writing the new sections

>>>>

\section{data is bad}
Now, one possible objection here is:
if those datasets are of such poor quality,
then maybe we just should not train any \ac{ML} model on these and should not use them to make automated decisions?
While this question is mostly beyond the scope of this document, let me offer some thoughts on this:
It is true that even after the application of de-biasing techniques,
the resulting models still should not be fully trusted,
but they can still ease the burden of checking everything manually;
similar to an email spam filter which is not perfect, but still very useful.
Or, put another way, it is always important to check what the realistic alternative is;
we should not compare a model to a non-existing perfect ideal, but to the actual solution that would be used instead.
One could imagine a hybrid approach where an automated system makes preliminary decisions,
but random samples are reviewed by humans and decisions can always be challenged.
Ideally, the model itself would tell us about decisions it is uncertain about.

Moreover, two of the three methods presented in this thesis require access to (unlabelled) \emph{unbiased} data,
which is used during the training process.
The remaining method requires summary statistics of the unbiased target dataset.
So, the criticism that we are only learning from biased data does not apply here.
This should allow us to be more confident in the predictions produced by those methods,
because the methods learn from additional, unbiased information.

\section{resume previous text}
